Structured Versus Decorated Cospans

### 2020-11-14 cospans.tex
a lot of work on section 4

We also give sufficient conditions for a decorated cospan double category to be equivalent to a structured cospan double category. 
Suppose $\A$ has finite colimits and $F \maps (\A , +) \to (\Cat, \times)$ is 
symmetric lax monoidal pseudofunctor.   Any such pseudofunctor gives an opfibration 
$R \maps \X \to \A$ where $\X = \int F$ is defined by the Grothendieck construction.
Let $\Rex$ be the 2-category of finitely cocomplete categories, finite-colimit-preserving
functors and natural transformations.  We show that if $F$ factors through $\mathbf{Rex}$ as
a pseudofunctor, the opfibration $R \maps \X \to \A$ is also a right adjoint.  From the accompanying left adjoint $L \maps \A \to \X$, we  construct a symmetric monoidal double category ${}_L \lCsp(\X)$ of structured cospans.  Finally, in \cref{Equiv} we prove that this structured cospan double category ${}_L \lCsp(\X)$ is equivalent to the decorated cospan double category $F \lCospan$.

@@

\section{Decorated cospans}\label{DecCospansDoubleCat}
In this section we build the symmetric monoidal double category $F \lCsp$ mentioned in the introduction. Then we obtain an underlying symmetric monoidal bicategory $F \mathbf{Csp}$ using a result of Shulman \cite{Shul} 
%!!!Christina: John suggested we might move all the horizontal bicategory and decategorified 1-category results together somewhere later/in the end.
and finally decategorify this to obtain a symmetric monoidal category $F\mathsf{Csp}$ which is a generalization of Fong's original decorated cospan category \cite{Fong}.  The definition of a lax monoidal pseudofunctor is recalled in \cref{sec:2cats}, and the basics of double categories are recalled in \cref{sec:doublecats}.

\begin{thm}\label{thm:decoratedcospans}
Let $\A$ be a category with chosen finite colimits and $F \maps \A \to \Cat$ a lax monoidal pseudofunctor. Then there exists a double category $F\lCsp$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical morphism is a morphism of $\A$,
\item a horizontal 1-cell is an $F$\define{-decorated cospan}: that is, 
a cospan in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
together with a \define{decoration} $x \in F(m)$,
\item a 2-morphism is a \define{map of} $F$\define{-decorated cospans}: that is, 
a map of cospans in $\A$:
\begin{equation}\label{eq:FCsp2morph}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$x \in F(m)$};
\node (F) at (3,-0.5) {$x' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[above] {$i'$} (D)
(C') edge node[above] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\end{equation}
together with a \define{morphism of decorations} $\tau \maps F(h)(x) \to x'$ in $F(m')$.
\end{itemize}
\end{thm}

@@

\begin{proof}
Let $f \maps c \to c'$ be a vertical 1-morphism in $F\lCsp$. We can lift $f$ to the companion horizontal 1-cell $\hat{f}$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$c$};
\node (B) at (1,0) {$c'$};
\node (C) at (2,0) {$c'$};
\node (D) at (1,-0.5) {$!_{c'} \in F(c')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(C) edge node[above]{$1$} (B);
\end{tikzpicture}
\]
and then obtain the following two 2-morphisms:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$c$};
\node (A') at (0,-0.5) {$c'$};
\node (B) at (1,0.5) {$c'$};
\node (C) at (2,0.5) {$c'$};
\node (C') at (2,-0.5) {$c'$};
\node (D) at (1,-0.5) {$c'$};
\node (E) at (3,0.5) {$!_{c'} \in F(c')$};
\node (F) at (3,-0.5) {$!_{c'} \in F(c')$};
\node (G) at (4,0.5) {$c$};
\node (H) at (5,0.5) {$c$};
\node (I) at (6,0.5) {$c$};
\node (G') at (4,-0.5) {$c$};
\node (H') at (5,-0.5) {$c'$};
\node (I') at (6,-0.5) {$c'$};
\node (J) at (7,0.5) {$!_{c} \in F(c)$};
\node (K) at (7,-0.5) {$!_{c'} \in F(c')$};
\node (L) at (1,-1) {$\tau_{1_{c'}} = 1_{!_{c'}}$};
\node (M) at (5,-1) {$\tau_f \maps F(f)(!_c) \to !_{c'}$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$f$} (B)
(C) edge node[above]{$1$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[left]{$1$} (C')
(A') edge node[above] {$1$} (D)
(C') edge node[above] {$1$} (D)
(B) edge node [left] {$1$} (D)
(G) edge node [above] {$1$} (H)
(G) edge node [left] {$1$} (G')
(H) edge node [left] {$f$} (H')
(G') edge node [above] {$f$} (H')
(I) edge node [above] {$1$} (H)
(I) edge node [left] {$f$} (I')
(I') edge node [above] {$1$} (H');
\end{tikzpicture}
\]
which satisfy the equations:
\[
\begin{tikzpicture}[scale=1.5]
\node (N) at (0,1.5) {$c$};
\node (O) at (1,1.5) {$c$};
\node (P) at (2,1.5) {$c$};
\node (Q) at (-1,1.5) {$!_c \in F(c)$};
\node (A) at (0,0.5) {$c$};
\node (A') at (0,-0.5) {$c'$};
\node (B) at (1,0.5) {$c'$};
\node (C) at (2,0.5) {$c'$};
\node (C') at (2,-0.5) {$c'$};
\node (D) at (1,-0.5) {$c'$};
\node (E) at (-1,0.5) {$!_{c'} \in F(c')$};
\node (F) at (-1,-0.5) {$!_{c'} \in F(c')$};
\node (G) at (4,1) {$c$};
\node (H) at (5,1) {$c$};
\node (I) at (6,1) {$c$};
\node (G') at (4,0) {$c'$};
\node (H') at (5,0) {$c'$};
\node (I') at (6,0) {$c'$};
\node (J) at (7,1) {$!_{c} \in F(c)$};
\node (K) at (7,0) {$!_{c'} \in F(c')$};
\node (Q) at (1,-1) {$\tau_f \maps F(f)(!_c) \to !_{c'}$};
\node (L) at (1,-1.5) {$\tau_{c'} = 1_{!_{c'}}$};
\node (M) at (5,-0.5) {$\tau_f \maps F(f)(!_c) \to !_{c'}$};
\node (R) at (3,0.5) {$=$};
\path[->,font=\scriptsize,>=angle 90]
(N) edge node[above]{$1$} (O)
(P) edge node[above]{$1$} (O)
(N) edge node[left]{$1$} (A)
(O) edge node[left]{$f$} (B)
(P) edge node[left]{$f$} (C)
(A) edge node[above]{$f$} (B)
(C) edge node[above]{$1$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[left]{$1$} (C')
(A') edge node[above] {$1$} (D)
(C') edge node[above] {$1$} (D)
(B) edge node [left] {$1$} (D)
(G) edge node [above] {$1$} (H)
(G) edge node [left] {$f$} (G')
(H) edge node [left] {$f$} (H')
(G') edge node [above] {$1$} (H')
(I) edge node [above] {$1$} (H)
(I) edge node [left] {$f$} (I')
(I') edge node [above] {$1$} (H');
\end{tikzpicture}
\]
\[
\begin{tikzpicture}[scale=1.5]
\node (G) at (-1,0.5) {$c$};
\node (H) at (-1,-0.5)  {$c'$};
\node (I) at (-2,0.5) {$c$};
\node (J) at (-2,-0.5) {$c$};
\node (A) at (0,0.5) {$c$};
\node (A') at (0,-0.5) {$c'$};
\node (B) at (1,0.5) {$c'$};
\node (C) at (2,0.5) {$c'$};
\node (C') at (2,-0.5) {$c'$};
\node (D) at (1,-0.5) {$c'$};
\node (E) at (1,1) {$!_{c'} \in F(c')$};
\node (F) at (1,-1) {$!_{c'} \in F(c')$};

(A) edge node[above]{$f$} (B)
(C) edge node[above]{$1$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[left]{$1$} (C')
(A') edge node[above] {$1$} (D)
(C') edge node[above] {$1$} (D)
(B) edge node [left] {$1$} (D)
(A) edge node[above]{$1$} (G)
(G) edge node[left]{$f$} (H)
(A') edge node[above]{$1$} (H)
(J) edge node[above] {$f$} (H)
(I) edge node[left] {$1$} (J)
(I) edge node [above] {$1$} (G);
\end{tikzpicture}
\]
The right hand sides of the above two equations are given respectively by the 2-morphisms $U_f$ and $1_{\hat{f}}$. The conjoint of $f$ is given by the $F$-decorated cospan $\check{f}$ which is just the opposite of the companion above:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$c'$};
\node (B) at (1,0) {$c'$};
\node (C) at (2,0) {$c$};
\node (D) at (4,0) {$!_{c'} \in F(c')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$1$} (B)
(C) edge node[above]{$f$} (B);
\end{tikzpicture}
\]
\end{proof}

The property of being fibrant is what allows us to lift the monoidal structure from the object category of a double category to its arrow category and obtain a symmetric monoidal bicategory. The following result, which only requires fibrancy on vertical 1-isomorphisms (isofibrancy) is due to Shulman \cite{Shul}:

\begin{lem}[Shulman]\label{Shul}
Let $\mathbb{X}$ be an isofibrant symmetric monoidal pseudo double category. Then the horizontal bicategory $H(\mathbb{X})$ of $\mathbb{X}$ is a symmetric monoidal bicategory which has:
\begin{enumerate}
\item{objects as those of $\mathbb{X}$,}
\item{morphisms as horizontal 1-cells of $\mathbb{X}$, and}
\item{2-morphisms as globular 2-morphisms of $\mathbb{X}$.}
\end{enumerate}
\end{lem}

\begin{thm}
Let $\A$ be a category with finite colimits and $F \maps (\A, +) \to \Cat$ a symmetric lax monoidal pseudofunctor. Then there exists a symmetric monoidal bicategory $F \mathbf{Csp} \mapseqq H(F\lCsp)$ which has:
\begin{enumerate}
\item{objects as those of $\A$,}
\item{morphisms as $F$-decorated cospans:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
and}
\item{2-morphisms as maps of cospans in $\A$ of the form:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0.5) {$c$};
\node (C) at (2,0) {$b$};
\node (E) at (1,-0.5) {$c'$};
\node (D) at (3,0.5) {$d \in F(c)$};
\node (F) at (3,-0.5) {$d' \in F(c')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[below]{$i'$} (E)
(B) edge node[left]{$h$} (E)
(C) edge node[below]{$o'$} (E);
\end{tikzpicture}
\]
together with a morphism $\tau \maps F(h)(d) \to d'$ in $F(c')$.}
\end{enumerate}
\end{thm}

\begin{proof}
This follows immediately by applying \cref{Shul} to the fibrant symmetric monoidal double category $F\lCsp$.
\end{proof}

This symmetric monoidal bicategory $F\mathbf{Csp}$ is a superior version of the symmetric monoidal bicategory $F$Cospan$(\A)$ constructed earlier by the second author \cite{Courser} in that there is greater flexibility in what 2-morphisms are allowed.

We can then decategorify this symmetric monoidal bicategory to obtain a symmetric monoidal category similar to the one obtained using Fong's result, but with larger isomorphism classes:

\begin{cor}
Let $\A$ be a category with finite colimits and $F \maps (\A, +) \to \Cat$ a symmetric lax monoidal pseudofunctor.  Then there exists a symmetric monoidal category $F\textnormal{Csp}$ which has:
\begin{enumerate}
\item{objects as those of $\A$ and}
\item{morphisms as isomorphism classes of $F$-decorated cospans of $\A$, where an $F$-decorated cospan is given by a pair:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d \in F(c)$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
Given another $F$-decorated cospan:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$c'$};
\node (C) at (2,0) {$b$};
\node (D) at (4,0) {$d' \in F(c')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[below]{$i'$} (B)
(C) edge node[below]{$o'$} (B);
\end{tikzpicture}
\]
these two $F$-decorated cospans are in the same isomorphism class if there exists an isomorphism $f \maps c \to c'$ such that following diagram commutes:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B') at (1,0.5) {$c$};
\node (B) at (1,-0.5) {$c'$};
\node (C) at (2,0) {$b$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[below]{$i'$} (B)
(C) edge node[below]{$o'$} (B)
(A) edge node[above]{$i$} (B')
(C) edge node[above]{$o$} (B')
(B') edge node[left]{$f$} (B);
\end{tikzpicture}
\]
and there exists an isomorphism $\tau \maps F(f)(d) \to d'$ in $F(c')$.}
\end{enumerate}
\end{cor}
In this symmetric monoidal category, isomorphism classes are as they should morally be, and the instance of two graphs having different edge sets does not prevent them from being in the same isomorphism class due to the isomorphism $\tau$.

@@

where $\kappa$ is the isomorphism $$\kappa \maps H(c_1 +_b c_2) \xrightarrow{\sim} H(c_1) +_{H(b)} H(c_2)$$ which comes from the finite colimit preserving functor $H \maps \A \to \mathsf{A'}$. The above diagram commutes by a similar argument to the one used in Theorem \cref{Equiv}. Similarly, for every object $c \in \A$, we have a unit comparison constraint $$\mathbb{H}_U \maps U_{\mathbb{H}(c)} \to \mathbb{H}(U_c)$$ given by the globular 2-isomorphism:

@@

The following theorem establishes the desired equivalence between these two different double categorical network frameworks of structured and decorated cospans in a formal way. We denote by $\bicat{Rex}$ the 2-category of finitely cocomplete categories and finitely cocontinuous functors, and the symmetric monoidal double structure of decorated cospans is worked out in \cref{thm:decoratedcospans,DC}.

\begin{thm}\label{Equiv}
Suppose $\ca{A}$ has finite colimits and $(F,\phi,\phi_0)\maps(\ca{A},+,0) \to (\bicat{Cat},\times,\mathbf{1})$ is a symmetric lax monoidal pseudofunctor that factors through $\bicat{Rex}$ as an ordinary pseudofunctor. 
Then the symmetric monoidal double category $F\lCsp$ of decorated cospans is equivalent to the symmetric monoidal double category $_L\lCsp(\inta F)$ of structured cospans, where $L \maps \A \to \inta F$ is the left adjoint of the induced Grothendieck opfibration $\U\maps \inta F \to \A$.
\end{thm}

@@

In \cite{MV}, the classical Grothendieck construction is generalized to the monoidal setting: lax monoidal structures on a pseudofunctor $F\maps\A\to\bicat{Cat}$ (called \emph{monoidal opindexed category}) bijectively correspond to monoidal structures on the total category $\inta F$ such that the induced opfibration is a strict monoidal functor and $\otimes_{\inta F}$ preserves cocartesian liftings (called \emph{monoidal fibration}). If the monoidal category $\A$ is in fact cocartesian, there is a further correspondence of the two structures with an ordinary pseudofunctor $F\maps\A\to\bicat{MonCat}$ into the 2-category of monoidal categories, strong monoidal functors and monoidal natural transformations.

\begin{thm}
 There is a 2-equivalence between the 2-categories of monoidal opfibrations and monoidal opindexed categories. If the base is cocartesian monoidal, there is a 2-equivalence between monoidal opfibrations and pseudofunctors into monoidal categories. 
\end{thm}

\begin{proof}
This is the content of \cite[Theorems~3.13\&4.1]{MV}. In summary, we have correspondences
\begin{gather}
\textrm{lax monoidal pseudofunctors }F\maps(\A,+)\to(\bicat{Cat},\times) \notag\\
\Updownarrow \notag\\
\textrm{monoidal opfibrations }U\maps(\X,\otimes)\to(\A,+) \label{monGroth}\\
\Updownarrow \notag\\
\textrm{pseudofunctors }F\maps\A\to\bicat{MonCat} \notag
\end{gather}
The second equivalence was also earlier observed by Shulman in \cite{Shul2}. In more detail, if $(\mu,\mu_0)$ is the lax monoidal structure of the pseudofunctor $F$ as recalled in \cref{sec:2cats}, each fibre category $\X_a=F(a)$ obtains a monoidal structure via
\begin{equation}\label{eq:explicitstructure1}
\otimes_a\maps F(a)\times F(a)\xrightarrow{\mu_{a,a}}F(a+a)\xrightarrow{F(\nabla)}F(a),\quad
I_a\maps\mathbf{1}\xrightarrow{\mu_0}F(0)\xrightarrow{F(!)}F(a)
\end{equation}
where $\nabla$ is the fold map. 
\end{proof}

These correspondences further restrict to the case when the Grothendieck category is specifically cocartesian monoidal itself. In that case, opfibrations
$(\X,+)\to(\A,+)$ that (strictly) preserve coproducts and the initial object (and such that $+$ preserves cocartesian liftings) bijectively correspond to pseudofunctors into the 2-category of cocartesian categories; for more details, see \cite[Corollary 4.7]{MV} and the related discussion. %!!! Christina: someone wrote Rex there, perhaps it was me. This is a subcategory, not all finite colimits are there.

@@

\begin{cor}\label{cor:fcocMonGroth}
 Suppose $\A$ is a finitely cocomplete category and $F\maps(\A,+,0)\to(\bicat{Cat},\times,1)$ is a lax monoidal pseudofunctor. If its corresponding pseudofunctor $\A\to\bicat{MonCat}$ via \cref{monGroth} 
 factors through $\bicat{Rex}$, then the Grothendieck category $\inta F$ has all finite colimits and the induced opfibration $\U_F\maps\inta F\to\A$ preserves them. 
\end{cor}

@@

Finally, if that pseudofunctor is lax monoidal $(F,\phi,\phi_0)\maps(\A,+,0)\to(\bicat{Cat},\times,\mathbf{1})$ to begin with, the monoidal Grothendieck construction in the cocartesian case discussed earlier \cref{eq:explicitstructure1} expresses $\bot_a$ as the image of the composite
\begin{equation}\label{eq:initialinfibre}
\mathbf{1}\xrightarrow{\phi_0}F(0_\A)\xrightarrow{F(u_a)}F(a) 
\end{equation}
\end{rmk}

@@

\begin{proof}[Proof of \cref{Equiv}]
Recall from \cref{thm:decoratedcospans} that the double category of decorated cospans $\dcsp{F}$ has objects and vertical 1-morphisms as in $\A$, while a horizontal 1-cell $a\tickar b$ is cospan $\cspn{a}{m}{b}{}{}$ in $\A$ decorated by an object $x\in F(m)$ and a 2-morphism is a map $k\maps m\to m'$ together with a morphism $(Fk)(x)\to x'$.

@@

Using a result of Shulman \cite{Shul}, each of the isofibrant symmetric monoidal double categories $F\lCsp$ and $_L \lCsp(\mathrm{X})$ give rise to underlying symmetric monoidal bicategories, namely $F\lCsp$ induces a symmetric monoidal bicategory $F\mathbf{Csp} \mapseqq H(F\lCsp)$ which has:
\begin{enumerate}
\item{objects as those of $\A$,}
\item{morphisms as horizontal 1-cells of $F\lCsp$, and}
\item{2-morphisms as globular 2-morphisms of $F\lCsp$.}
\end{enumerate} 
Likewise, $_L \lCsp(\X)$ induces a symmetric monoidal bicategory $H( {_L \lCsp(\X)})$ which has:
\begin{enumerate}
\item{objects as those of $\A$,}
\item{morphisms as horizontal 1-cells of $_L \lCsp(\X)$, and}
\item{2-morphisms as globular 2-morphisms of $_L \lCsp(\X)$.}
\end{enumerate}
Another result of Shulman \cite{Shul2} is the following:
\begin{lem}[Shulman, Prop.\ B.3]
An equivalence of fibrant double categories induces a biequivalence of horizontal bicategories.
\end{lem}

@@

>>>>

We also give sufficient conditions for a decorated cospan double category to be equivalent to a structured cospan double category. 
Suppose $\A$ has finite colimits and $F \maps (\A , +) \to (\Cat, \times)$ is 
symmetric lax monoidal pseudofunctor.   Any such pseudofunctor gives an opfibration 
$R \maps \X \to \A$ where $\X = \int F$ is defined by the Grothendieck construction.
Let $\Rex$ be the 2-category of finitely cocomplete categories, finite-colimit-preserving
functors and natural transformations.  We show that if $F$ factors through $\mathbf{Rex}$ as
a pseudofunctor, the opfibration $R \maps \X \to \A$ is also a right adjoint.  From the accompanying left adjoint $L \maps \A \to \X$, we  construct a symmetric monoidal double category ${}_L \lCsp(\X)$ of structured cospans.  Finally, in \cref{thm:equiv} we prove that this structured cospan double category ${}_L \lCsp(\X)$ is equivalent to the decorated cospan double category $F \lCospan$.

@@

\section{Decorated cospans}\label{DecCospansDoubleCat}

In this section we build the symmetric monoidal double category $F \lCsp$ mentioned in the introduction, and then study the functoriality of this construction.   The definition of a lax monoidal pseudofunctor is recalled in \cref{sec:2cats}, and the basics of double categories are recalled in \cref{sec:doublecats}.

\begin{thm}\label{thm:decorated_cospans}
Let $\A$ be a category with chosen finite colimits and $F \maps \A \to \Cat$ a lax monoidal pseudofunctor. Then there exists a double category $F\lCsp$ in which
\begin{itemize}
\item an object is an object of $\A$,
\item a vertical morphism is a morphism of $\A$,
\item a horizontal 1-cell is an $F$\define{-decorated cospan}: that is, 
a cospan in $\A$:
\[
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0) {$a$};
\node (B) at (1,0) {$m$};
\node (C) at (2,0) {$b,$};
%\node (D) at (2.6,0) {$x \in F(m))$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B);
\end{tikzpicture}
\]
together with a \define{decoration} $x \in F(m)$,
\item a 2-morphism is a \define{map of} $F$\define{-decorated cospans}: that is, 
a map of cospans in $\A$:
\begin{equation}\label{eq:FCsp2morph}
\begin{tikzpicture}[scale=1.5]
\node (A) at (0,0.5) {$a$};
\node (A') at (0,-0.5) {$a'$};
\node (B) at (1,0.5) {$m$};
\node (C) at (2,0.5) {$b$};
\node (C') at (2,-0.5) {$b'$};
\node (D) at (1,-0.5) {$m'$};
\node (E) at (3,0.5) {$x \in F(m)$};
\node (F) at (3,-0.5) {$x' \in F(m')$};
\path[->,font=\scriptsize,>=angle 90]
(A) edge node[above]{$i$} (B)
(C) edge node[above]{$o$} (B)
(A) edge node[left]{$f$} (A')
(C) edge node[right]{$g$} (C')
(A') edge node[above] {$i'$} (D)
(C') edge node[above] {$o'$} (D)
(B) edge node [left] {$h$} (D);
\end{tikzpicture}
\end{equation}
together with a \define{morphism of decorations} $\tau \maps F(h)(x) \to x'$ in $F(m')$.
\end{itemize}
\end{thm}

@@

@@

where $\kappa$ is the isomorphism $$\kappa \maps H(c_1 +_b c_2) \xrightarrow{\sim} H(c_1) +_{H(b)} H(c_2)$$ which comes from the finite colimit preserving functor $H \maps \A \to \mathsf{A'}$. The above diagram commutes by a similar argument to the one used in \cref{thm:equiv}. Similarly, for every object $c \in \A$, we have a unit comparison constraint $$\mathbb{H}_U \maps U_{\mathbb{H}(c)} \to \mathbb{H}(U_c)$$ given by the globular 2-isomorphism:

@@

The following theorem establishes the desired equivalence between these two different double categorical network frameworks of structured and decorated cospans in a formal way. We denote by $\bicat{Rex}$ the 2-category of finitely cocomplete categories and finitely cocontinuous functors, and the symmetric monoidal double structure of decorated cospans is worked out in \cref{thm:decorated_cospans,DC}.

\begin{thm} \label{thm:equiv}
Suppose $\ca{A}$ has finite colimits and $(F,\phi,\phi_0)\maps(\ca{A},+) \to (\bicat{Cat},\times,1)$ is a symmetric lax monoidal pseudofunctor that factors through $\bicat{Rex}$ as an ordinary pseudofunctor. 
Then the symmetric monoidal double category $F\lCsp$ of decorated cospans is equivalent to the symmetric monoidal double category $_L\lCsp(\inta F)$ of structured cospans, where $L \maps \A \to \inta F$ is the left adjoint of the induced Grothendieck opfibration $\U\maps \inta F \to \A$.
\end{thm}

@@

In \cite{MV}, the classical Grothendieck construction is generalized to the monoidal setting: lax monoidal structures on a pseudofunctor $F\maps\A\to\bicat{Cat}$ (called \emph{monoidal opindexed categories}) bijectively correspond to monoidal structures on the total category $\inta F$ such that the induced opfibration is a strict monoidal functor and $\otimes_{\inta F}$ preserves cocartesian liftings (called \emph{monoidal fibrations}). If the monoidal category $\A$ is in fact cocartesian, there is a further correspondence of the two structures with an ordinary pseudofunctor $F\maps\A\to\bicat{MonCat}$ into the 2-category of monoidal categories, strong monoidal functors and monoidal natural transformations.

\begin{lem}
 There is a 2-equivalence between the 2-categories of monoidal opfibrations and monoidal opindexed categories. If the base is cocartesian monoidal, there is a 2-equivalence between monoidal opfibrations and pseudofunctors into monoidal categories. 
\end{lem}

\begin{proof}
This was shown by Moeller and the third author \cite[Theorems~3.13\&4.1]{MV}. In summary, we have correspondences
\begin{gather}
\textrm{lax monoidal pseudofunctors }F\maps(\A,+)\to(\bicat{Cat},\times) \notag\\
\Updownarrow \notag\\
\textrm{monoidal opfibrations }U\maps(\X,\otimes)\to(\A,+) \label{monGroth}\\
\Updownarrow \notag\\
\textrm{pseudofunctors }F\maps\A\to\bicat{MonCat} \notag
\end{gather}
The second equivalence was also earlier observed by Shulman \cite{Shul2008}. In more detail, if $(\mu,\mu_0)$ is the lax monoidal structure of the pseudofunctor $F$ as recalled in \cref{sec:2cats}, each fibre category $\X_a=F(a)$ obtains a monoidal structure via
\begin{equation}\label{eq:explicitstructure1}
\otimes_a\maps F(a)\times F(a)\xrightarrow{\mu_{a,a}}F(a+a)\xrightarrow{F(\nabla)}F(a),\quad
I_a\maps\mathbf{1}\xrightarrow{\mu_0}F(0)\xrightarrow{F(!)}F(a)
\end{equation}
where $\nabla$ is the fold map. 
\end{proof}

These correspondences further restrict to the case when the Grothendieck category is specifically cocartesian monoidal itself. In that case, opfibrations
$(\X,+)\to(\A,+)$ that strictly preserve chosen coproducts and initial object, and such that $+$ preserves cocartesian liftings, bijectively correspond to pseudofunctors into the 2-category of cocartesian categories; for more details, see \cite[Corollary 4.7]{MV} and the related discussion. %!!! Christina: someone wrote Rex there, perhaps it was me. This is a subcategory, not all finite colimits are there.

@@

\begin{cor}\label{cor:fcocMonGroth}
 Suppose $\A$ is a finitely cocomplete category and $F\maps(\A,+)\to(\bicat{Cat},\times)$ is a lax monoidal pseudofunctor. If its corresponding pseudofunctor $\A\to\bicat{MonCat}$ via \cref{monGroth} 
 factors through $\bicat{Rex}$, then the Grothendieck category $\inta F$ has all finite colimits and the induced opfibration $\U_F\maps\inta F\to\A$ preserves them. 
\end{cor}

@@

Finally, if that pseudofunctor is lax monoidal $(F,\phi,\phi_0)\maps(\A,+)\to(\bicat{Cat},\times)$ to begin with, the monoidal Grothendieck construction in the cocartesian case discussed earlier \cref{eq:explicitstructure1} expresses $\bot_a$ as the image of the composite
\begin{equation}\label{eq:initialinfibre}
\mathbf{1}\xrightarrow{\phi_0}F(0_\A)\xrightarrow{F(u_a)}F(a) 
\end{equation}
\end{rmk}

@@

\begin{proof}[Proof of \cref{thm:equiv}]
Recall from \cref{thm:decorated_cospans} that the double category of decorated cospans $\dcsp{F}$ has objects and vertical 1-morphisms as in $\A$, while a horizontal 1-cell $a\tickar b$ is cospan $\cspn{a}{m}{b}{}{}$ in $\A$ decorated by an object $x\in F(m)$ and a 2-morphism is a map $k\maps m\to m'$ together with a morphism $(Fk)(x)\to x'$.

@@

@@

\section{Bicategorical and categorical aspects}
\label{spinoffs}

While double categories are a natural context for studying cospans, bicategories are more 
familiar---and of course, mere \emph{categories} are even more so!   Luckily, all our results 
phrased in the language of double categories have analogues for bicategories and categories.  
We explain those here.

As noted for example by Shulman \cite{Shulman2010}, any double category $\lX$ gives rise to a bicategory $\bX$ in which:
\begin{itemize}
\item objects are objects of $\lX$,
\item morphisms are horizontal 1-cells of $\lX$,
\item 2-morphisms are \define{globular} 2-morphisms of $\lX$, meaning 2-morphisms whose source and target vertical 1-morphisms are identities,
\item composition of morphisms is given by horizontal composition of horizontal 1-cells in $\lX$,
\item vertical and horizontal composition of 2-morphisms are given by vertical and horizontal
composition of 2-morphisms in $\lX$.
\end{itemize}
Furthermore, any bicategory $\bX$ whatsoever gives rise to a category $\X$ in which:
\begin{itemize}
\item objects are objects of $\bX$,
\item morphisms are isomorphism classes of morphisms of $\bX$.   
\end{itemize}
Thus, the double category $F\lCsp$ of structured cospans constructed in \cref{thm:decorated_cospans} automatically gives rise to a bicategory, which we call $F\bCsp$, and a category, which we call $F\Csp$.

In \cref{DC} we gave conditions under which the double category $F\lCsp$ becomes symmetric
monoidal.   We would like the bicategory $F\bCsp$ and the category $F\Csp$ to become symmetric
monoidal under the same conditions, and indeed this is true.  Shulman \cite[Thm.\ 1.2]{Shulman2010} showed that when $\lX$ is `fibrant', any symmetrical monoidal structure on $\lX$ gives one on $\bX$.     We can apply this to decorated cospans as follows:
